---
title: "SQL"
output: html_document
---

##Filtering Columns and Rows

In SQL one can filter columns and rows of a selected table. Columns can be filtered using the SELECT command and rows can be filtered using the WHERE command. Following is an example of this using the Lahman baseball database.

Before the tables can be filtered, the Lahman and sqldf packages need to be loaded.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Here is an example of filters used in the Batting table of the Lahman database to look at the home run totals of the 1927 Yankees.

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

One can also get more selective with the data, using the same filters. Following are some more specific questions within the Batting table of the Lahman database, the queries used to find the answers, and the answers themselves.

####Examples

Question 1: Find all instances where Yankees have hit 40 or more home runs.

For this question, one will not only want to see the number of homeruns and the team, but most likely the player and the year as well. Additionally, instead of using an equals sign in the WHERE clause, one can use greater or less than signs as well to narrow or broaden searches. 

Here is the query line used to find the answer to Question 1.

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

Question 2: Find an instance where a player had more than 40 homeruns but less than 60 strikeouts.

```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO from Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

Question 3: Find all instances of Phillies in the 1970s hitting more than 30 homeruns.

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='PHI'and yearID>1970 and yearID<=1979 and HR>30"
sqldf(query)
```

##Ordering Data Outputs

One can also order output. To order the information, a ORDER BY clause is used. This will default to ascending order (smallest to largest). To organize in descending order, the DESC command will be placed after the ORDER BY command.

####Examples

Question 1: Find every instance of a player hitting more than 50 homeruns, ordered from highest to lowest homerun total.

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

Question 2: Find all instances of a player striking out less than 10 times. Make sure each player has at least 400 at-bats (AB). Order by least to most strikeouts.

```{r}
query<-"SELECT playerID,yearID,teamID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

##Aggregation

Aggregation allows one to group and condense information in different ways, essentially 'squashing' different peces of data together. This is done by using the GROUP BY command. When using GROUP BY, one must then specify how the other columns asked for should be grouped. Should they be averaged, should they be added, or does one only want to see the minimum or maximum? These modifiers are placed on the other column names.This is how one would specify what should be done to those data sets:

* Summation: sum(colname)

* Average: avg(colname)

* Maximum: max(colname)

* Minimum: min(colname)

If one wants to continue filtering after they have grouped, they can use the HAVING command, which functions as the filtering WHERE command when used after GROUP BY. ORDER BY can be used in the same way as before, coming at the end of the command list.

####Examples

Question 1: Find the career homerun totals for Babe Ruth.

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

Question 2: Find career homerun totals for all players but list only those with a total of 600 or more, put in descending order.

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```

Question 3: What players have the highest average per season homerun totals? List only those with averages greater than 30, and order from highest to lowest.

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```

##Joining

Joining allows tables to be joined together based on common data to increase the data output options. This is done through the INNER JOIN and ON commands. INNER JOIN indicates which tables are being joined together. ON indicates the common data that the tables will be joined by. If the common data appears alone in the code, it will need to either have specified which of the joined tables it is coming from, or replaced by the inforamtion from another table. In the first example, playerID in the SELECT phrase was replaced by nameFIRST and nameLAST to make reading the table easier.

####Examples

Question 1: First and last name, along with team, year, and homeruns for Babe Ruth. 

```{r}
query<-"SELECT nameFIRST,nameLAST,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

Question 2: Find all instances of players getting more than 50 homeruns starting with the greatest number. Include first and last names, team, year, and home runs.

```{r}
query<-"SELECT nameFIRST,nameLAST,teamID,yearID,HR
FROM Batting INNER JOIN Master
On Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

Question 3: Find Babe Ruth's homeruns by season, with the team name displayed.

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"
sqldf(query)
```

Question 4: Find Babe Ruth's homerun totals, with first and last name and the full team name displayed.

```{r}
query<-"SELECT nameFIRST,nameLAST,name,Batting.yearID,Batting.HR 
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams 
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

Question 5: Find career homerun totals for players who have at least 600 homeruns. Display from the highest amount to the lowest, with the first and last names.

```{r}
query<-"SELECT nameFIRST,nameLAST,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```

Question 6: List all players, first and last name, with the first name of Bob who averaged more than one million dollars salary over their career.

```{r}
query<-"SELECT nameFIRST,nameLAST,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
GROUP BY Salaries.playerID
HAVING nameFIRST='Bob' and avg(salary)>1000000
ORDER BY avg(salary) DESC"
sqldf(query)
```

